---
title: "Appendix B: Replicator Dynamic Phase Diagram Generator Code (R)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r packages}
# packages used
library(EvolutionaryGames)
library(rgl)
```

### Mouse, Hawk, Retaliator, Prober-Retaliator (MS&P's payoffs)
```{r}
# create the payoff matrix for the game
A <- matrix(c(29, 19.5, 29, 17.2,
              80, -19.5, -18.1, -18.9,
              29, -22.3, 29, 23.1, 
              56.7, -20.1, 26.9, 21.9), 
            4, byrow=TRUE)

# specify the names of the strategies, in order of column
strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

# specify the initial population state
state <- c(.25, .25, .25, .25)

# plot phase diagram (3D for 4 strategies)
phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
# if you run with noRGL = F, the rgl package will create an interactive 3D diagram
# that you can rotate! 
```

### Mouse, Hawk, Retaliator (MS&P)
```{r}
A <- matrix(c(29, 19.5, 29, 
              80, -19.5, -18.1,
              29, -22.3, 29), 
            3, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator")

state <- matrix(c(.33, .33, .34), 1, 3, byrow = T)

# this one is 3D for 2 strategies
# it also has a contour map showing the speed of change 
# and a vector field showing the direction
phaseDiagram3S(A, dynamic = Replicator, params = NULL, trajectories = state, 
               contour = T, vectorField = T, strategies)

# you can also find the set of ESSs and graph their paths
ESset(A, strategies)
```

### Mouse, Hawk, Prober-Retaliator (MS&P)
```{r}
A <- matrix(c(29, 19.5, 17.2,
              80, -19.5, -18.9,
              56.7, -20.1, 21.9), 
            3, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Prober-Retaliator")

ESset(A, strategies)

state <- matrix(c(.34, .33, .33), 1, 3, byrow = T)

phaseDiagram3S(A, dynamic = Replicator, params = NULL, trajectories = state, 
               contour = F, vectorField = F, strategies)
```

### Mouse, Retaliator, Prober-Retaliator (MS&P)
This one has no ESS. 
```{r}
A <- matrix(c(29, 29, 17.2,
              29, 29, 23.1, 
              56.7, 26.9, 21.9), 
            3, byrow=TRUE)

strategies <- c("Mouse", "Retaliator", "Prober-Retaliator")

state <- matrix(c(.34, .33, .33), 1, 3, byrow = T)

phaseDiagram3S(A, dynamic = Replicator, params = NULL, trajectories = state, 
               contour = F, vectorField = F, strategies)
```

### Hawk, Retaliator, Prober-Retaliator (MS&P)
```{r}
A <- matrix(c(-19.5, -18.1, -18.9,
              -22.3, 29, 23.1, 
              -20.1, 26.9, 21.9), 
            3, byrow=TRUE)

strategies <- c("Hawk", "Retaliator", "Prober-Retaliator")

ESset(A, strategies)
```

# Phase Diagrams from My Computer Simulation

### Mouse, Hawk, Retaliator, Prober-Retaliator (my payoffs)
```{r}
A <- matrix(c(30, 19.75, 30, 15.3,
              79.75, -20.1, -15.2, -14.9,
              30, -25.3, 30, 12.2, 
              67.6, -25.7, 14.9, 3.1), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.25, 0.25, 0.25, 0.25)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```


### Mouse, Hawk, Retaliator, Prober-Retaliator (my, Mouse starts at 0)
```{r}
state <- c(0, 0.33, 0.33, 0.33)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```

### Mouse, Hawk, Retaliator, Prober-Retaliator (my, Mouse starts at 0.01)
```{r}
state <- c(0.01, 0.33, 0.33, 0.33)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```

### Mouse, Hawk, Retaliator, Prober-Retaliator (my, Mouse starts at 0.04)
```{r}
state <- c(0.04, 0.32, 0.32, 0.32)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```


# Phase Diagrams from Changing Model Parameters

### Increasing the probability of serious injury (p_i = 0.90)
```{r}
A <- matrix(c(30, 19.75, 30, 15.4,
              79.75, -10.1, 63.4, 59.8,
              30, -83.9, 30, -69.5, 
              67.6, -80.3, 53.8, -11.7), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.25, 0.25, 0.25, 0.25)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```

### Increasing the probability of probing behavior (p_p = 0.5)
```{r}
A <- matrix(c(30, 19.75, 30, 19.2,
              79.75, -20.2, -14.8, -15.3,
              30, -25.6, 30, -23.5, 
              79.2, -25.1, -13.8, -20.4), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.25, 0.25, 0.25, 0.25)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```


### Increasing the cost of serious injury (c = 500)
```{r}
A <- matrix(c(30, 19.75, 30, 15.4,
              79.75, -220.2, -205.8, -205.3,
              30, -234.6, 30, -53.4, 
              67.6, -235.1, -42.5, -97.2), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.1, 0.7, 0.1, 0.1)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```

### Increasing the payoff of winning (v = 120)
```{r}
A <- matrix(c(60, 19.75, 60, 19.2,
              139.75, 10, 15, 14,
              60, 5.5, 60, 41.3, 
              123.9, 5.4, 46.2, 34), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.25, 0.25, 0.25, 0.25)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```

### Decreasing the probability of retaliation (p_r = 0.5)
```{r}
A <- matrix(c(30, 19.75, 30, 15.4,
              79.75, -20.2, 14.5, -8.6,
              30, -54.4, 30, -3.2, 
              67.6, -32.4, 21.5, -2), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.25, 0.25, 0.25, 0.25)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```

### Increasing the maximum payoff of saving time and energy, producing longer contests (u_init = 40)
```{r}
A <- matrix(c(30, 39.75, 30, 30.9,
              99.75, -10.2, 29.9, 4.2,
              30, -51.2, 30, -23.6, 
              89.9, -24.9, 24.2, -8.8), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.25, 0.25, 0.25, 0.25)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```

### Decreasing the maximum payoff of saving time and energy, producing shorter contests (u_init = 5)
```{r}
A <- matrix(c(30, 4.75, 30, 19.1,
              64.75, -27.6, 18.8, -2.1,
              30, -50.4, 30, 19.8, 
              43.1, -37.1, 27.7, 18.8), 
            4, byrow=TRUE)

strategies <- c("Mouse", "Hawk", "Retaliator", "Prober-Retaliator")

state <- c(0.25, 0.25, 0.25, 0.25)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)

state <- c(0.1, 0.1, 0.4, 0.4)

phaseDiagram4S(A, Replicator, NULL, state, noRGL = T, strategies)
```
